TestEvol: A Tool for Analyzing Test-Suite Evolution

### 2012-11-02 icse2013demo-testevol/paper.tex
Revised abstract

\documentclass[conference]{IEEEtran}
\synctex=1

\usepackage{times}
\usepackage{graphicx}
\usepackage{epsfig}
% \usepackage{url}
% \usepackage{hyperref}
\usepackage[bookmarks=false]{hyperref}
\usepackage{xspace}
\usepackage{float}
\usepackage{latexsym}
\usepackage{natbib}
\usepackage[usenames]{color}
\usepackage{multirow}
\usepackage{listings}
\usepackage{textcomp}

@@

\author{\IEEEauthorblockN{XXX}
\IEEEauthorblockA{
% College of Computing\\
Georgia Institute of Technology\\
% Atlanta, USA\\
\{orso\}@cc.gatech.edu}
}

@@

\note{This is a VERY rough draft that I put together by copying and
  pasting from the FSE paper. I left some specific TODOs in, but the
  paper really needs a major pass.}

\todo{Restructure the content to make sure we comply with what is
  requested from a demo paper: Tool-based demonstrations describe
  novel aspects of early prototypes or mature tools. The tool
  demonstrations must communicate clearly the following information to
  the audience:\\
  - the envisioned users;\\
  - the software engineering challenge it proposes to address;\\
  - the methodology it implies for its users; and\\
  - the results of validation studies already conducted for mature tools, or the design of planned studies for early prototypes.}

\todo{Reduce the description of the approach and extend the
  description of the implementation. Leandro, can you add some details
  there?}

\todo{Add snapshots.}

\begin{abstract}
  Test suites, once created, rarely remain static. Just like the
  application they are testing, they evolve throughout their lifetime.
  Test obsolescence is probably the most known reason for test-suite
  evolution---test cases cease to work because of changes in the code
  and must be suitably repaired. Repairing existing test cases
  manually, however, can be extremely time consuming, especially for
  large test suites, which has motivated the recent development of
  automated test-repair techniques. We believe that, for developing
  effective repair techniques that are applicable in real-world
  scenarios, a fundamental prerequisite is a thorough understanding of
  how test cases evolve in practice. Without such knowledge, we risk
  to develop techniques that may work well for only a small number of
  tests or, worse, that may not work at all in most realistic cases.
  Unfortunately, to date there are no studies in the literature that
  investigate how test suites evolve. To tackle this problem, in this
  demo we present \tool, a tool that enables the systematic study of
  test-suite evolution for Java programs and JUnit test cases. We also
  present examples of applications of \tool\ to real-world programs
  and their unit test suites that show the usefulness and practical
  applicability of the tool.
\end{abstract}

@@

Test cases form the first line of defense against the introduction of
software faults, especially when retesting modified software.  With
the availability of convenient testing frameworks, such as JUnit, and
the adoption of agile development methodologies, in particular,
writing unit test cases is an increasingly common practice nowadays.
Developers routinely create test suites that consist of unit tests and
that they run periodically on their code.

Test suites are not static entities: they constantly evolve along with
the application they test.  In particular, changes in the application
can break test cases---in some cases, even a small change in the
application code can affect a large number of tests. In general, a
test failure on a new version of the software can either expose
application faults or result from a problem with the test itself.  On
observing a failure, the first task for the developer is to determine
which of these two is the cause of the failure and, once this
distinction is made, either fix the problem in the code or fix the
broken~test.

A broken test, if it covers a valid functionality, should ideally be
repaired. Alternatively, if the repair is unduly complex to perform or
if the test was designed to cover a functionality that no longer
exists in the application, the test should be removed from the test
suite. To illustrate with an example,
Figure~\ref{fig:pmd-javatokenizertest} shows two versions of a unit
test case from the test suite of \subject{PMD}, one of the programs
used in our demo. A change in \subject{PMD}'s API broke the original
version of the test case, which had to be fixed by adding a call to
method \lang{SourceCode.readSource} and removing one parameter from
the call to method \lang{Tokenizer.tokenize} (lines 6 and 7 in
Figure~\ref{fig:pmd-javatokenizertest}(b), respectively).

@@

To the best of our knowledge, to date there are no studies in the
literature that investigate how unit test suites evolve. To address
this issue, we defined a technique~\cite{pinto12} that combines
various static- and dynamic-analysis techniques to compute the
differences between the test suites associated with two versions of a
program and categorize such changes along two dimensions: the static
differences between the tests in the two test suites and the
behavioral differences between such tests.  We implemented our
technique in a tool, \tool, which enables a systematic study of
test-suite evolution. \tool\ targets Java programs and JUnit test
suites, as Java is a widely used language and JUnit is the de-facto
standard unit-testing framework for Java.  Given two versions of a
program and its test suite, \tool\ automatically computes differences
in the behavior of the test suites on the two program versions,
classifies the actual repairs performed between the versions, and
computes the coverage attained by the tests on the two program
versions.

@@

\section{Our Approach for Studying Test-Suite Evolution}
\label{sec:test-evolution}
\subsection{Definitions and Terminology}
\label{sec:terminology}

@@

\begin{figure}[t]
\centering
\small
\tabcolsep=3pt
\textbf{(a) Test $t$ exists in $S$ and $S'$ and is modified}
\\ [2pt]
\begin{tabular}{|l||l|}
\hline
%\multicolumn{1}{|c||}{} & \multicolumn{1}{|c}{} &
%\multicolumn{1}{|c|}{$t$ is not} \\
%\multicolumn{1}{|c||}{} & \multicolumn{1}{|c|}{$t$ is modified}\\
%\hline \hline
$\testfunc{P'}{t} = \fail \; \wedge$ &
$t$ is repaired \\
$\testfunc{P'}{t'} = \pass$ &
\multicolumn{1}{r|}{\scriptsize [\catrep{}]}\\
\hline
\multirow{3}{28mm}{$\testfunc{P'}{t} = \pass \; \wedge \testfunc{P'}{t'} = \pass$} &
$t$ is refactored, updated to test a different \\
& scenario, or is made more/less discriminating \\
& \multicolumn{1}{r|}{\scriptsize [\catref{}]} \\
\hline
\end{tabular}
\\ [8pt]
%
\textbf{(b) Test $t$ is removed in $S'$}
\\ [2pt]
\begin{tabular}{|l||l|}
\hline
\multirow{2}{*}{$\testfunc{P'}{t} = \failre \, | \, \failae$} &
$t$ is too difficult to fix \\
& \multicolumn{1}{r|}{\scriptsize [\catdelaere{}]} \\
\hline
\multirow{2}{*}{$\testfunc{P'}{t} = \failce$} & $t$ is obsolete or is
too difficult to fix \\
& \multicolumn{1}{r|}{\scriptsize [\catdelce{}]} \\
\hline
\multirow{2}{*}{$\testfunc{P'}{t} = \pass$} & $t$ is redundant \\
& \multicolumn{1}{r|}{\scriptsize [\catdelp{}]} \\
\hline 
\end{tabular}
\\ [8pt]
%
\textbf{(c) Test $t'$ is added in $S'$}
\\ [2pt]
\begin{tabular}{|l||l|}
  \hline
  \multirow{2}{*}{$\testfunc{P}{t'} = \failre \, | \, \failae$} &
  $t'$ is added to validate a bug fix \\
  & \multicolumn{1}{r|}{\scriptsize [\cataddaere{}]} \\
  \hline
  \multirow{3}{*}{$\testfunc{P}{t'} = \failce$} &
  $t'$ is added to test a new functionality \\
  & or a code refactoring \\
  & \multicolumn{1}{r|}{\scriptsize [\cataddce{}]} \\
  \hline
  \multirow{3}{*}{$\testfunc{P}{t'} = \pass$} &
  %% TODO: Check: $t'$ is added to test a new functionality \\
  $t'$ is added to test an existing feature\\
  & or for coverage-based augmentation \\
  & \multicolumn{1}{r|}{\scriptsize [\cataddp{}]} \\
  \hline
\end{tabular}
\caption{Scenarios considered in our investigation. Given two system
  versions $S = (P, T)$ and $S' = (P', T')$, the three scenarios are:
  (a) $t$ exists in $T$ and $T'$ and is modified, (b) $t$ exists in
  $T$ but not in $T'$, (c) $t'$ exists in $T'$ but not in $T$.}
\vspace*{-5pt}
\label{fig:study-design}
\end{figure}

\subsubsection{Category \catrep{}: Repaired Tests}
\label{sec:category-rep}

The \catrep{} category corresponds to the case where $t$ is repaired
so that, after the modifications, it passes on $P'.$ As discussed in
the Introduction, Figure~\ref{fig:pmd-javatokenizertest} shows an
example of such a test repair. The code fragments in
Listings~\ref{c1-old} and~\ref{c1-new} present another example of
repair that involves a simpler code modification than the one in
Figure~\ref{fig:pmd-javatokenizertest}. The example is taken from
\subject{Gson}, one of the programs used in our demo, and involves a
test case from \subject{Gson} version~2.0 that was fixed in the
subsequent version~2.1. Test \lang{testNullField} had to be fixed
because constructor \lang{FieldAttributes(Class<?> declClazz, Field
  f)} from version~2.0 was modified in version~2.1 to take only one
parameter (of type \lang{Field}).
%% was no longer available on version 2.1. In the new version of the
%% \lang{FieldAttributes} class, the constructor receives only one
%% parameter of type \lang{Field}.

\vspace*{-10pt}
\lstinputlisting[language=java,label=c1-old,caption=\small{Unit test
  for class FieldAttributes} (Gson v2.0),
captionpos=t,frame=tb]{code/OldTestRep.java}

\vspace*{-5pt}
\lstinputlisting[language=java,label=c1-new,caption=\small{Unit test
  for class FieldAttributes} (Gson v2.1),
captionpos=t,belowcaptionskip=4pt,frame=tb]{code/TestRep.java}

@@

A test repair may involve multiple such changes. For example, the
repair shown in Figure~\ref{fig:pmd-javatokenizertest} involves the
addition of a method call~(in~line~6) and the deletion of a method
parameter (in line~7). The repair illustrated in Listings~\ref{c1-old}
and~\ref{c1-new} also involves the deletion of a method parameter.
%
For \textit{assertion changes}, we consider cases in which an
assertion is added, an assertion is deleted, the expected value of an
assertion is modified, or the assertion is modified but the expected
value is unchanged.
%
Finally, we currently group together \textit{data-value changes} and
\textit{control-flow changes}.

The rationale underlying our classification scheme is that different
classes of changes may require different types of repair analyses.
Although this is not necessarily the case, at least the search
strategy for candidate repairs would differ for the different
categories of changes. Consider the case of method-parameter deletion,
for instance, for which one could attempt a repair by simply deleting
some of the actual parameters. Whether this repair would work depends
on the situation. For the code in
Figure~\ref{fig:pmd-javatokenizertest}, for example, it would not work
because the deletion of one of the parameters in the call to
\lang{tokenize()} is insufficient by itself to fix the test---a new
method call (to \lang{readSource()}) has to be added as well for the
test to work correctly.  Similarly, for the case of method-parameter
addition, one could conceivably attempt straightforward fixes by
constructing equivalence classes for the new parameters and selecting
a value from each equivalence class (\eg positive, negative, and zero
values for an integer parameter). In our study, we found cases where
such an approach would in fact repair broken tests. In this case too,
however, such a solution is in general not enough.

\subsubsection{Category \catref{}: Refactored Tests}
\label{sec:category-c2}

The \catref{} category captures scenarios in which a test $t$ is
modified in $S'$ even though $t$ passes on $P'$.
Listings~\ref{c2-old} and~\ref{c2-new} show an example of one such
change from \subject{Commons Math}. Unit test \lang{testDistance} was
refactored to invoke method \lang{sqrt()} in class \lang{FastMath}, a
newly added class in the new release, instead of the same method in
\lang{java.lang.Math}.  There are different reasons why changes in
this category might occur.

\vspace{-8pt}
\lstinputlisting[language=java,label=c2-old,caption=\small{Unit test
  from class Vector3DTest} (Commons Math v2.1),
captionpos=t,frame=tb]{code/OldTestRef.java}

\vspace{-8pt}
\lstinputlisting[language=java,label=c2-new,caption=\small{Unit test
  from class Vector3DTest} (Commons Math v2.2),
captionpos=t,frame=tb]{code/TestRef.java}

@@

\subsubsection{Category \catdelaere{}: Hard-To-Fix Tests}
\label{sec:category-delaere}

@@

This category includes tests that fail on $P'$ with a runtime
exception or an assertion violation. These may be instances where the
tests should have been fixed, as the functionality that they test in
$P$ still exists in $P'$, but the tests were discarded instead. One
plausible hypothesis is that tests in this category involve repairs of
undue complexity, for which the investigation of new repair techniques
to aid the developer might be particularly useful.  We performed a
preliminary investigation of this hypothesis by manually examining ten
randomly selected tests in this category and found that all the
examined tests were in fact obsolete.

\subsubsection{Category \catdelce: Obsolete Tests}
\label{sec:category-delce}

A test that fails with a compilation error on the new program version
is obsolete because of API changes. Listing~\ref{c4-example}
illustrates a test in this category taken from \subject{JodaTime}.
This test was deleted because the tested method
\lang{Chronology.getBuddhist()} was removed in the subsequent version
of \subject{JodaTime}.

% \vspace{-4pt}
\lstinputlisting[language=java,label=c4-example,caption=\small{Unit
  test from class TestChronology} (JodaTime v2.0),frame=tb]
{code/TestDel-Ce.java}
%This method was
%already marked as deprecated in the previous version.

Although for this category of deletion too, one could postulate that
the tests were removed because they were too difficult to fix, we
believe this not to be the case in most practical occurrences.
Instead, the more likely explanation is that the tests were removed
simply because the tested methods were no longer present. Also in this
case, we investigated our hypothesis by manually examining ten
randomly selected cases. Indeed, our manual investigation confirmed
that, for the cases we analyzed, the tested functionality was either
removed or provided through alternative methods, which requires the
development of new tests rather than fixes to the existing ones.

\subsubsection{Category \catdelp: Redundant Tests}
\label{sec:category-delp}

This category includes tests that are removed even though they pass on
$P'$.  Listing~\ref{c5-example} illustrates an example of one such
test, taken from \subject{Commons Lang}. In the new version, this test
was replaced by a more sophisticated one, shown in
Listing~\ref{c5.2-example}.

\begin{figure}[h]
\lstinputlisting[language=java,label=c5-example,caption=\small{Unit
  test from class MatrixIndexExceptionTest} (Commons Lang
v2.1),frame=tb] {code/TestDel-P.java}
\vspace{4pt}
\end{figure}

\lstinputlisting[language=java,label=c5.2-example,caption=\small{Unit
  test implemented to replace the removed one} (Commons Lang
v2.2),frame=tb] {code/TestDel-PII.java}

@@

\subsubsection{Category \cataddaere: Bug-Fix Tests}
\label{sec:category-addaere}

This category includes added tests that fail on $P$ with a runtime
exception or an assertion violation. In this case, the functionality
that the added test $t'$ was designed to test exists in $P$ but is not
working as expected (most likely because of a fault). The program
modifications between $P$ and $P'$ would ostensibly have been made to
fix the fault, which causes $t'$ to pass on $P'$. Thus, $t'$ is added
to the test suite to validate the bug fix. Listing~\ref{c6-example}
illustrates a test that fits this profile, as it was added to
\subject{Commons Math} version 2.1 to validate a bug fix ({\small
  \url{https://issues.apache.org/jira/browse/MATH-391}}). In the
previous version of \lang{ArrayRealVector}, the creation of a
zero-length vector results in a runtime exception.

% \begin{figure}[h]
\lstinputlisting[language=java,label=c6-example,caption=\small{\mbox{Unit
  test from class ArrayRealVectorTest (Commons Math v2.2)}},
captionpos=t,frame=tb]{code/TestAdd-AeRe.java}
% \end{figure}

\subsubsection{Category \cataddce: New-Features Tests}
\label{sec:category-addce}

The added tests in this category fail on $P$ with a compilation error,
which indicates that the API accessed by the tests does not exist in
$P$. Thus, the added test $t'$ is created to test new code in $P'$,
where the code could have been added as part of a refactoring or, more
likely, to add new functionality.  Listing~\ref{c7-example}
illustrates a test from \subject{JFreeChart} that covers a new
functionality: method \lang{getMinY()} did not exist in previous
versions of class \lang{TimeSeries}.

\lstinputlisting[language=java,label=c7-example,caption=\small{Unit
  test from class TimeSeriesTests} (JFreeChart v1.0.14),
captionpos=t,frame=tb]{code/TestAdd-Ce.java}

\subsubsection{Category \cataddp: Coverage-Augmentation Tests}
\label{sec:category-addp}

This category considers cases where the added test $t'$ passes on
$P$. Clearly, $t'$ would have been a valid test in the old system as
well. One would expect that the addition of $t'$ increases program
coverage (\ie $\covfunc{P'}{T'} \supset \covfunc{P'}{T' -
  \{t'\}}$). Moreover, if $t'$ covers different statements in $P$ and
$P'$ (assuming that there is a way of matching statements between $P$
and $P'$), the plausible explanation is that $t'$ was added to test
the changes made between $P$ and $P'$. However, if $t'$ covers the
same statements in both program versions, it would have been added
purely to increase code coverage (and not to test any added or
modified code).  Listing~\ref{c8-example} illustrates an added test
case that increases code coverage: the new test method
\lang{testFindRangeBounds} covers the case where the parameter of
method \lang{findRangeBounds()} is null.

\lstinputlisting[language=java,label=c8-example,caption=\small{Unit test
from class XYErrorRendererTests} (JFreeChart v1.0.14),frame=tb]
{code/TestAdd-P.java}

@@

\bibliographystyle{abbrv}
\bibliography{paper}

>>>>

\documentclass[conference]{IEEEtran}
%\synctex=1

\usepackage{times}
\usepackage{graphicx}
\usepackage{epsfig}
% \usepackage{url}
% \usepackage{hyperref}
\usepackage[bookmarks=false]{hyperref}
\usepackage{xspace}\usepackage{float}
\usepackage{latexsym}
\usepackage{natbib}
\usepackage[usenames]{color}
\usepackage{multirow}
\usepackage{listings}
\usepackage{textcomp}

@@

\author{
\IEEEauthorblockN{Leandro Sales Pinto}
\IEEEauthorblockA{
Politecnico di Milano \\
pinto@elet.polimi.it}
\and
\IEEEauthorblockN{Saurabh Sinha}
\IEEEauthorblockA{
IBM Research -- India \\
saurabhsinha@in.ibm.com}
\and
\IEEEauthorblockN{Alessandro Orso}
\IEEEauthorblockA{
Georgia Institute of Technology \\
orso@cc.gatech.edu}
}

@@

% \note{This is a VERY rough draft that I put together by copying and
%   pasting from the FSE paper. I left some specific TODOs in, but the
%   paper really needs a major pass.}
% \todo{Restructure the content to make sure we comply with what is
%   requested from a demo paper: Tool-based demonstrations describe
%   novel aspects of early prototypes or mature tools. The tool
%   demonstrations must communicate clearly the following information to
%   the audience:\\
%   - the envisioned users;\\
%   - the software engineering challenge it proposes to address;\\
%   - the methodology it implies for its users; and\\
%   - the results of validation studies already conducted for mature tools, or the design of planned studies for early prototypes.}
% \todo{Reduce the description of the approach and extend the
%   description of the implementation. Leandro, can you add some details
%   there?}
% \todo{Add snapshots.}

\begin{abstract}
  Test suites, once created, rarely remain static. Just like the
  application they are testing, they evolve throughout their lifetime.
  Test obsolescence is probably the most known reason for test-suite
  evolution---test cases cease to work because of changes in the code
  and must be suitably repaired. There are several reasons why it is
  important to achieve a thorough understanding of how test cases
  evolve in practice. In particular, researchers who develop automated
  test-repair techniques---an increasingly active research area---can
  use such understanding to develop more effective repair techniques
  that are applicable in real-world scenarios. More generally,
  analyzing test-suite evolution can help testers better understand
  how test cases are modified during maintenance and improve the test
  evolution process, an extremely time consuming activity for any non
  trivial test suite.  Unfortunately, there are no existing tools that
  facilitate investigation of test-suite evolution. To tackle this
  problem, we developed \tool, a tool that enables the systematic
  study of test-suite evolution for Java programs and JUnit test
  cases.  This demo presents \tool and demonstrates its usefulness and
  practical applicability by showing how \tool\ can be used on
  real-world software and unit test suites.
\end{abstract}

@@

Test suites are not static entities: they constantly evolve along with
the application they test. For instance, new tests can be added to
test new functionality and existing tests can be refactored, repaired,
or deleted.  In particular, changes in the application can break test
cases---in some cases, even a small change in the application code can
affect a large number of tests.

A broken test, if it covers a valid functionality, should ideally be
repaired. Alternatively, if the repair is unduly complex to perform or
if the test was designed to cover a functionality that no longer
exists in the application, the test should be removed from the test
suite. To illustrate with an example,
Figure~\ref{fig:pmd-javatokenizertest} shows two versions of a unit
test case from the test suite of \subject{PMD}, one of the programs we
analyzed in our previous work~\cite{pinto12}. A change in
\subject{PMD}'s API broke the original version of the test case, which
had to be fixed by adding a call to method
\lang{SourceCode.readSource} and removing one parameter from the call
to method \lang{Tokenizer.tokenize} (lines 6 and 7 in
Figure~\ref{fig:pmd-javatokenizertest}(b), respectively).

@@

To the best of our knowledge, to date there are no studies in the
literature that investigate how unit test suites evolve. To address
this issue, we defined a technique~\cite{pinto12} that combines
various static- and dynamic-analysis techniques to compute the
differences between the test suites associated with two versions of a
program and categorize such changes along two dimensions: the static
differences between the tests in the two test suites and the
behavioral differences between such tests.

We implemented our technique in a tool, \tool, which enables a
systematic study of test-suite evolution. \tool\ targets Java programs
and JUnit test suites, as Java is a widely used language and JUnit is
the de-facto standard unit-testing framework for Java.  Given two
versions of a program and its test suite, \tool\ automatically
computes differences in the behavior of the test suites on the two
program versions, classifies the actual repairs performed between the
versions, and computes the coverage attained by the tests on the two
program versions.

@@

\section{Studying Test-Suite Evolution}
\label{sec:test-evolution}

\subsection{Definitions and Terminology}
\label{sec:terminology}

@@

\begin{figure}[t]
\centering
\footnotesize
\tabcolsep=3pt
\textbf{(a) Test $t$ exists in $S$ and $S'$ and is modified}
\\ [2pt]
\begin{tabular}{|l||l|}
\hline
%\multicolumn{1}{|c||}{} & \multicolumn{1}{|c}{} &
%\multicolumn{1}{|c|}{$t$ is not} \\
%\multicolumn{1}{|c||}{} & \multicolumn{1}{|c|}{$t$ is modified}\\
%\hline \hline
$\testfunc{P'}{t} = \fail \; \wedge$ &
$t$ is repaired \\
$\testfunc{P'}{t'} = \pass$ &
\multicolumn{1}{r|}{\scriptsize [\catrep{}]}\\
\hline
\multirow{3}{28mm}{$\testfunc{P'}{t} = \pass \; \wedge \testfunc{P'}{t'} = \pass$} &
$t$ is refactored, updated to test a different \\
& scenario, or is made more/less discriminating \\
& \multicolumn{1}{r|}{\scriptsize [\catref{}]} \\
\hline
\end{tabular}
\\ [8pt]
%
\textbf{(b) Test $t$ is removed in $S'$}
\\ [2pt]
\begin{tabular}{|l||l|}
\hline
\multirow{2}{*}{$\testfunc{P'}{t} = \failre \, | \, \failae$} &
$t$ is too difficult to fix \\
& \multicolumn{1}{r|}{\scriptsize [\catdelaere{}]} \\
\hline
\multirow{2}{*}{$\testfunc{P'}{t} = \failce$} & $t$ is obsolete or is
too difficult to fix \\
& \multicolumn{1}{r|}{\scriptsize [\catdelce{}]} \\
\hline
\multirow{2}{*}{$\testfunc{P'}{t} = \pass$} & $t$ is redundant \\
& \multicolumn{1}{r|}{\scriptsize [\catdelp{}]} \\
\hline 
\end{tabular}
\\ [8pt]
%
\textbf{(c) Test $t'$ is added in $S'$}
\\ [2pt]
\begin{tabular}{|l||l|}
  \hline
  \multirow{2}{*}{$\testfunc{P}{t'} = \failre \, | \, \failae$} &
  $t'$ is added to validate a bug fix \\
  & \multicolumn{1}{r|}{\scriptsize [\cataddaere{}]} \\
  \hline
  \multirow{3}{*}{$\testfunc{P}{t'} = \failce$} &
  $t'$ is added to test a new functionality \\
  & or a code refactoring \\
  & \multicolumn{1}{r|}{\scriptsize [\cataddce{}]} \\
  \hline
  \multirow{3}{*}{$\testfunc{P}{t'} = \pass$} &
  %% TODO: Check: $t'$ is added to test a new functionality \\
  $t'$ is added to test an existing feature\\
  & or for coverage-based augmentation \\
  & \multicolumn{1}{r|}{\scriptsize [\cataddp{}]} \\
  \hline
\end{tabular}
\caption{Scenarios considered in our investigation. Given two system
  versions $S = (P, T)$ and $S' = (P', T')$, the three scenarios are:
  (a) $t$ exists in $T$ and $T'$ and is modified, (b) $t$ exists in
  $T$ but not in $T'$, (c) $t'$ exists in $T'$ but not in $T$.}
\vspace*{-5pt}
\label{fig:study-design}
\end{figure}

\subsubsection{Category \catrep{} (Repaired Tests)}
\label{sec:category-rep}

The \catrep{} category corresponds to the case where $t$ is repaired
so that, after the modifications, it passes on $P'.$ As discussed in
the Introduction, Figure~\ref{fig:pmd-javatokenizertest} shows an
example of such a test repair.

%% The code fragments in Listings~\ref{c1-old} and~\ref{c1-new} present
%% another example of repair that involves a simpler code modification
%% than the one in Figure~\ref{fig:pmd-javatokenizertest}. The example is
%% taken from \subject{Gson}, one of the programs used in our demo, and
%% involves a test case from \subject{Gson} version~2.0 that was fixed in
%% the subsequent version~2.1. Test \lang{testNullField} had to be fixed
%% because constructor \lang{FieldAttributes(Class<?> declClazz, Field
%%   f)} from version~2.0 was modified in version~2.1 to take only one
%% parameter (of type \lang{Field}).
%% %% was no longer available on version 2.1. In the new version of the
%% %% \lang{FieldAttributes} class, the constructor receives only one
%% %% parameter of type \lang{Field}.

%% \vspace*{-10pt}
%% \lstinputlisting[language=java,label=c1-old,caption=\small{Unit test
%%   for class FieldAttributes} (Gson v2.0),
%% captionpos=t,frame=tb]{code/OldTestRep.java}

%% \vspace*{-5pt}
%% \lstinputlisting[language=java,label=c1-new,caption=\small{Unit test
%%   for class FieldAttributes} (Gson v2.1),
%% captionpos=t,belowcaptionskip=4pt,frame=tb]{code/TestRep.java}

@@

A test repair may involve multiple such changes. For example, the
repair shown in Figure~\ref{fig:pmd-javatokenizertest} involves the
addition of a method call (line~6) and the deletion of a method
parameter (line~7).
%% The repair illustrated in Listings~\ref{c1-old} and~\ref{c1-new} also
%% involves the deletion of a method parameter.
%
For \textit{assertion changes}, we consider cases in which an
assertion is added, an assertion is deleted, the expected value of an
assertion is modified, or the assertion is modified but the expected
value is unchanged.
%
Finally, we currently group together \textit{data-value changes} and
\textit{control-flow changes}.

The rationale underlying our classification scheme is that different
classes of changes may require different types of repair analyses.
Although this is not necessarily the case, at least the search
strategy for candidate repairs would differ for the different
categories of changes. Consider the case of method-parameter deletion,
for instance, for which one could attempt a repair by simply deleting
some of the actual parameters. Whether this repair would work depends
on the situation. For the code in
Figure~\ref{fig:pmd-javatokenizertest}, for example, it would not work
because the deletion of one of the parameters in the call to
\lang{tokenize()} is insufficient by itself to fix the test---a new
method call (to \lang{readSource()}) has to be added as well for the
test to work correctly.
%% Similarly, for the case of method-parameter addition, one could
%% conceivably attempt straightforward fixes by constructing equivalence
%% classes for the new parameters and selecting a value from each
%% equivalence class (\eg positive, negative, and zero values for an
%% integer parameter).  In our study, we found cases where such an
%% approach would in fact repair broken tests. In this case too, however,
%% such a solution is in general not enough.

\subsubsection{Category \catref{} (Refactored Tests)}
\label{sec:category-c2}

The \catref{} category captures scenarios in which a test $t$ is
modified in $S'$ even though $t$ passes on $P'$.
%% Listings~\ref{c2-old} and~\ref{c2-new} show an example of one such
%% change from \subject{Commons Math}. Unit test \lang{testDistance} was
%% refactored to invoke method \lang{sqrt()} in class \lang{FastMath}, a
%% newly added class in the new release, instead of the same method in
%% \lang{java.lang.Math}.

%% \vspace{-5pt}
%% \lstinputlisting[language=java,label=c2-old,caption=\small{Unit test
%%   from class Vector3DTest} (Commons Math v2.1),
%% captionpos=t,frame=tb]{code/OldTestRef.java}

%% \vspace{-5pt}
%% \lstinputlisting[language=java,label=c2-new,caption=\small{Unit test
%%   from class Vector3DTest} (Commons Math v2.2),
%% captionpos=t,frame=tb]{code/TestRef.java}

@@

\subsubsection{Category \catdelaere{} (Hard-To-Fix Tests)}
\label{sec:category-delaere}

@@

This category includes tests that fail on $P'$ with a runtime
exception or an assertion violation. These may be instances where the
tests should have been fixed, as the functionality that they test in
$P$ still exists in $P'$, but the tests were discarded instead. One
plausible hypothesis is that tests in this category involve repairs of
undue complexity, for which the investigation of new repair techniques
to aid the developer might be particularly useful.
%% We performed a preliminary investigation of this hypothesis by
%% manually examining ten randomly selected tests in this category and
%% found that all the examined tests were in fact obsolete.

\subsubsection{Category \catdelce (Obsolete Tests)}
\label{sec:category-delce}

A test that fails with a compilation error on the new program version
is obsolete because of API changes.
%
%% Listing~\ref{c4-example} illustrates a test in this category taken
%% from \subject{JodaTime}.  This test was deleted because the tested
%% method \lang{Chronology.getBuddhist()} was removed in the subsequent
%% version of \subject{JodaTime}.
%
%% % \vspace{-4pt}
%% \lstinputlisting[language=java,label=c4-example,caption=\small{Unit
%%   test from class TestChronology} (JodaTime v2.0),frame=tb]
%% {code/TestDel-Ce.java}
%% %This method was
%% %already marked as deprecated in the previous version.
%
Although for this category of deletion too, one could postulate that
the tests were removed because they were too difficult to fix, we
believe this not to be the case in most practical occurrences.
Instead, the more likely explanation is that the tests were removed
simply because the tested methods were no longer present.
%% Also in this case, we investigated our hypothesis by manually
%% examining ten randomly selected cases. Indeed, our manual
%% investigation confirmed that, for the cases we analyzed, the tested
%% functionality was either removed or provided through alternative
%% methods, which requires the development of new tests rather than fixes
%% to the existing ones.

\subsubsection{Category \catdelp: (Redundant Tests)}
\label{sec:category-delp}

This category includes tests that are removed even though they pass on
$P'$. These tests would typically be redundant, by some measure of
redundancy (\eg code-coverage based). In some cases, a test may be
replaced with a different one.

%% Listing~\ref{c5-example} illustrates an example of one such test,
%% taken from \subject{Commons Lang}. In the new version, this test was
%% replaced by a more sophisticated one, shown in
%% Listing~\ref{c5.2-example}.

%% \begin{figure}[h]
%% \lstinputlisting[language=java,label=c5-example,caption=\small{Unit
%%   test from class MatrixIndexExceptionTest} (Commons Lang
%% v2.1),frame=tb] {code/TestDel-P.java}
%% \vspace{4pt}
%% \end{figure}

%% \lstinputlisting[language=java,label=c5.2-example,caption=\small{Unit
%%   test implemented to replace the removed one} (Commons Lang
%% v2.2),frame=tb] {code/TestDel-PII.java}

@@

\subsubsection{Category \cataddaere: (Bug-Fix Tests)}
\label{sec:category-addaere}

This category includes added tests that fail on $P$ with a runtime
exception or an assertion violation. In this case, the functionality
that the added test $t'$ was designed to test exists in $P$ but is not
working as expected (most likely because of a fault). The program
modifications between $P$ and $P'$ would ostensibly have been made to
fix the fault, which causes $t'$ to pass on $P'$. Thus, $t'$ is added
to the test suite to validate the bug fix.

%% Listing~\ref{c6-example} illustrates a test that fits this profile, as
%% it was added to \subject{Commons Math} version 2.1 to validate a bug
%% fix ({\small
%%   \url{https://issues.apache.org/jira/browse/MATH-391}}). In the
%% previous version of \lang{ArrayRealVector}, the creation of a
%% zero-length vector results in a runtime exception.

%% % \begin{figure}[h]
%% \lstinputlisting[language=java,label=c6-example,caption=\small{\mbox{Unit
%%   test from class ArrayRealVectorTest (Commons Math v2.2)}},
%% captionpos=t,frame=tb]{code/TestAdd-AeRe.java}
%% % \end{figure}

\subsubsection{Category \cataddce (New-Features Tests)}
\label{sec:category-addce}

The added tests in this category fail on $P$ with a compilation error,
which indicates that the API accessed by the tests does not exist in
$P$. Thus, the added test $t'$ is created to test new code in $P'$,
where the code could have been added as part of a refactoring or, more
likely, to add new functionality.

%% Listing~\ref{c7-example} illustrates a test from \subject{JFreeChart}
%% that covers a new functionality: method \lang{getMinY()} did not exist
%% in previous versions of class \lang{TimeSeries}.

%% \lstinputlisting[language=java,label=c7-example,caption=\small{Unit
%%   test from class TimeSeriesTests} (JFreeChart v1.0.14),
%% captionpos=t,frame=tb]{code/TestAdd-Ce.java}

\subsubsection{Category \cataddp (Coverage-Augmentation Tests)}
\label{sec:category-addp}

This category considers cases where the added test $t'$ passes on
$P$. Clearly, $t'$ would have been a valid test in the old system as
well. One would expect that the addition of $t'$ increases program
coverage (\ie $\covfunc{P'}{T'} \supset \covfunc{P'}{T' -
  \{t'\}}$). Moreover, if $t'$ covers different statements in $P$ and
$P'$ (assuming that there is a way of matching statements between $P$
and $P'$), the plausible explanation is that $t'$ was added to test
the changes made between $P$ and $P'$. However, if $t'$ covers the
same statements in both program versions, it would have been added
purely to increase code coverage (and not to test any added or
modified code).

%% Listing~\ref{c8-example} illustrates an added test case that increases
%% code coverage: the new test method \lang{testFindRangeBounds} covers
%% the case where the parameter of method \lang{findRangeBounds()} is
%% null.

%% \lstinputlisting[language=java,label=c8-example,caption=\small{Unit test
%% from class XYErrorRendererTests} (JFreeChart v1.0.14),frame=tb]
%% {code/TestAdd-P.java}

@@

\bibliographystyle{abbrv}
{\footnotesize
\bibliography{paper}
}
